Introduction to LLL\\``Cryptography''

### 2021-05-27 paper.tex
Init RSA Attack

\chapter{Applications}

\section{Attack Knapsack}

\section{Attack RSA}

>>>>

\chapter{Cryptanlysis}

% \section{Attack Knapsack}

\section{RSA introduction}

\textbf{RSA} is one of the earliest and most used asymmetric cryptosystem.
The usual step to generate a public/private key for \textbf{RSA} is the following

\begin{enumerate}
    \item Fix $e = 65537$ or $e = 3$ (public).
    \item Find two primes $p, q$ such that $p - 1$ and $q - 1$ are relatively prime to $e$, i.e. gcd$(e, p-1) = 1$ and gcd$(e, q-1) = 1$.
    \item Compute $N = p * q$ and $\phi(n) = (p-1) * (q-1)$
    \item Calculate $d$ (private) as the multiplicative inverse of $e$ modulo $\phi(n)$.
    \item $(N, e)$ is the public key, $(N, d)$ is the private key.
\end{enumerate}

To encrypt a message $m$ with \textbf{textbook RSA}

\begin{center}
    $c = m^e \mod N$
\end{center}

To decrypt a ciphertext $c$

\begin{center}
    $m = c^d \mod N$
\end{center}

\section{Lattices against RSA}

It's easy to find the roots of a univariate polynomial over the integers. Finding the roots of \textbf{modular} polynomial is hard, example:

\begin{center}
    $f(x) \equiv 0 \mod N$
\end{center}

Suppose $N$ is an \textbf{RSA} modulus and we don't know the factorization of it. Let's have an univariate integer polynomial $f(x)$ with degree $n$

\begin{center}
    $f(x) = x^n + a_{n-1}x^{n-1} + a_{n-2}x^{n-2} + \cdots + a_1n + a_0$
\end{center}

Coppersmith showed how we can recover the value $x_0$ such that $f(x_0) \equiv 0 \mod N$, with $x_0 < N^{\frac{1}{n}}$ in polynomial time
using the following theorem

\begin{theorem}
    \textbf{Howgrave-Graham}
\end{theorem}

\textit{Let $g(x)$ be an univariate polynomial with n monomials and $m$ be a positive integer.
    If we have some restraint $X$ and the following equations hold}

\begin{center}
    \begin{eqnarray}
         g(x_0) \equiv 0 \mod N^m, |x_0| \le X \\
         \lVert g(xX) \rVert < \frac{N}{\sqrt{n}} \label{eq:pol}
    \end{eqnarray}
\end{center}

Then $g(x_0) = 0$ holds over the integers.

\subsection{Example}

We have a 100-bit \textbf{RSA} modulus

\begin{center}
    $N=0xf046522fb555a90bdc558fc93$ and $e = 3$.
\end{center}

Before the encryption the message $m$ is padded as

\[
    z = pad || m = 0x74686973206b65793a00 || m
\]

where $||$ is the concatenation.\\

The ciphertext is

\[
    c = z^e \mod N = 0x8185d7de1d7cb8801e69c0bf
\]

Suppose that we don't know the factorization of $N$ and we would like to know the message $m$.
However we know the padding and the length of $m < 2^{16}$.

\vspace*{10px}

Let's define

\[
a = 0x74686973206b65793a0000
.\]

which is the known padding string.

\vspace*{10px}

Thus we have that $c = (a + m)^3 \mod N$, for an unkown small $m$.
We can define $f(x) = (a + x)^3 - c$, and so we setup the problem to find a small root $m$ such that $f(m) \equiv 0 \mod N$

\begin{align*}
    f(x) = x^3 + 0x15d393c596142306bae0000x^2 + 0x1b53c5e184a49b39f9ad9eedbx \\
    + 0x9bb23cefd94fac9cc9cee9943
\end{align*}

\vspace*{10px}

\textbf{Lattice contruction}. Let the coefficients of $f$ be $f(x) = x^3 + f_2x^2 + f_1x + f_0$
and $X = 2^{16}$ be the upper bound of the size of the root $m$. We can construct the matrix

\[
B = 
\begin{pmatrix}
    X^3 & f_2X^2 & f_1X & f_0 \\
    0 & NX^2 & 0 & 0 \\
    0 & 0 & NX & 0 \\
    0 & 0 & 0 & N \\
\end{pmatrix}
\]

If we apply LLL to the lattice we get

\clearpage
We can use lattices to solves the problem:

\begin{itemize}
    \item If we have some polynomial which share the same root $x_0$ over $N^m$ we can represent each of them as a row from lattice,
        and every linear integer combination of rows between them will yield another polynomial having a root $x_0$.
    \item Using \textbf{LLL} on the lattice we can find a reduced lattice basis, such that the norm of the shortest vector will
        satisfy the inequality of \ref{eq:pol}. \textbf{LLL} will only use \textbf{linear integer operations} on the basis vectors,
        in this way we preserve the root $x_0$ and the bound $X$.
    \item Define the shortest vector in the reduced basis as $v = (v_0, v_1, \ldots, v_n)$. We can construct the polynomial
        \begin{center}
            \[
                g(x) = v_0 + \frac{\displaystyle v_1}{X}x + \frac{v_2}{X^2}x^2 + \cdots + \frac{v_n}{X^n}x^n
            \]
        \end{center}
\end{itemize}
Since $g(x)$ is on the lattice we know that

\begin{center}
    \begin{eqnarray}
        g(x_0) \equiv 0 \mod N^m \\
        \vert x_0| \le X \\
        \lVert g(xX) \rVert < \frac{N^m}{\sqrt{deg(g) + 1}}
    \end{eqnarray}
\end{center}

\section{Key recovery ECDSA}
